---
title: "Hands-on Exercise 4a: Visualising Distribution"
author: "Htet Kay Khaing"
date: "May 2, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 4.1 Learning outcomes
In this chapter, we will explore two newer methods for visualising distributions, namely the **ridgeline plot** and **raincloud plot**, using **ggplot2** and its extensions. These build on earlier techniques introduced in Chapter 1, such as the histogram, density plot, boxplot, notch plot, and violin plot.

## 4.2 Getting Started
### .2.1 Installing and loading the packages
For the purpose of this exercise, the following R packages will be used, they are:

- ggridges, a ggplot2 extension specially designed for plotting ridgeline plots,
- ggdist, a ggplot2 extension specially design for visualising distribution and uncertainty,
- tidyverse, a family of R packages to meet the modern data science and visual communication needs,
- ggthemes, a ggplot extension that provides the user additional themes, scales, and geoms for the ggplots package, and
- colorspace, an R package provides a broad toolbox for selecting individual colors or color palettes, manipulating these colors, and employing them in various kinds of visualisations.

The code chunk below will be used load these R packages into RStudio environment.

```{r}
pacman::p_load(ggdist, ggridges, ggthemes,
               colorspace, tidyverse)
```

### 4.2.2 Data import
For the purpose of this exercise, *Exam_data.csv* will be used.

In the code chunk below, [`read_csv()`](https://readr.tidyverse.org/reference/read_delim.html) of [**readr**](https://readr.tidyverse.org/) package is used to import *Exam_data.csv* into R and saved it into a tibble data.frame.

```{r}
exam <- read_csv("data/Exam_data.csv")
```









