---
title: "Take-home Exercise 1"
author: "Htet Kay Khaing"
date: "May 7, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 1. Overview/Background

Motor insurance portfolios usually contain many policies with no claims and a small number of policies with severe claim outcomes. Because of this, aggregate profitability can look acceptable even when selected segments are loss-making or highly volatile. This study uses visual analytics to examine the distribution of key variables and to identify portfolio factors that may influence profitability and volatility.

## 2. The tasks

1.  **Apply suitable univariate visual analytics techniques** to explore the distributions and characteristics of numerical and categorical variables.
2.  **Apply suitable bivariate visual analytics techniques** to investigate and statistically validate factors influencing the profitability and volatility of the insurance portfolio.

For this report, **profitability** is measured mainly using `underwriting_result`, `underwriting_margin` and `loss_ratio`. **Volatility** is examined using claim occurrence, claim frequency, the spread of loss ratio and upper-tail loss ratio behaviour.

::: {.callout-note}
The analysis is exploratory. The statistical tests validate whether observed differences are unlikely to be random, but they do not prove direct causality. Business interpretation is still required because the dataset is observational.
:::

## 3. Getting started

### 3.1 Installing and loading the packages

For this exercise, the following R packages will be used:

-   **tidyverse** for data wrangling and ggplot2 visualisation,
-   **scales** for formatting values,
-   **patchwork** for combining plots,
-   **ggdist** for raincloud-style distribution plots,
-   **ggridges** for ridgeline plots, and
-   **knitr** for report formatting support.

The following code chunk uses `p_load()` of **pacman** package to check if tidyverse packages are installed in the computer. If they are, the libraries will be called into R.

```{r}
pacman::p_load(tidyverse, scales, patchwork,
               ggdist, ggridges, knitr)
```

## 4. The Data

### 4.1 Data import

The data file used is [Dataset of motor insurance portfolio.csv](https://data.mendeley.com/datasets/sw4jmdb2sm/1). The file is semicolon-delimited, so `read_delim()` is used instead of `read_csv()`.

```{r}
motor_raw <- read_delim(
  "data/Dataset of motor insurance portfolio.csv",
  delim = ";",
  show_col_types = FALSE
)
# glimpse(motor_raw)
```

### 4.2 Data overview

```{r, echo=FALSE}
overview_tbl <- tibble(
  metric = c(
    "Number of records",
    "Number of variables",
    "Unique insured IDs",
    "Years covered",
    "Total premium",
    "Total incurred claims",
    "Total claims",
    "Total exposure"
  ),
  value = c(
    comma(nrow(motor_raw)),
    comma(ncol(motor_raw)),
    comma(n_distinct(motor_raw$insured_id)),
    paste(sort(unique(motor_raw$year)), collapse = ", "),
    dollar(sum(motor_raw$total_premium, na.rm = TRUE), prefix = "$", accuracy = 1),
    dollar(sum(motor_raw$total_incurred, na.rm = TRUE), prefix = "$", accuracy = 1),
    comma(sum(motor_raw$total_claims, na.rm = TRUE)),
    comma(sum(motor_raw$total_exposure, na.rm = TRUE), accuracy = 0.1)
  )
)

knitr::kable(
  overview_tbl,
  caption = "Dataset overview: Motor insurance portfolio transaction records"
)
```

The dataset has **354,140** records and **47** variables. It covers policy transaction records from **2022, 2023, 2024**, which allows both cross-sectional and time-based portfolio analysis.

### 4.3 Data Cleaning

::: {.panel-tabset}
## Check Missing Values
```{r, echo=FALSE}
missing_wide <- summarise(
  motor_raw,
  across(everything(), ~sum(is.na(.x)))
)

missing_tbl <- pivot_longer(
  missing_wide,
  cols = everything(),
  names_to = "variable",
  values_to = "missing_records"
)

missing_tbl <- mutate(
  missing_tbl,
  missing_pct = missing_records / nrow(motor_raw)
)

missing_tbl <- filter(
  missing_tbl,
  missing_records > 0
)

missing_tbl <- arrange(
  missing_tbl,
  desc(missing_records)
)

missing_display_tbl <- mutate(
  missing_tbl,
  missing_records = comma(missing_records),
  missing_pct = percent(missing_pct, accuracy = 0.01)
)

knitr::kable(
  missing_display_tbl,
  caption = "Variables with missing values"
)
```

```{r, echo=FALSE}
missing_plot_tbl <- mutate(
  missing_tbl,
  variable = fct_reorder(variable, missing_pct)
)

ggplot(
  missing_plot_tbl,
  aes(x = missing_pct, y = variable)
) +
  geom_col(fill = "#e75480") +
  geom_text(
    aes(label = percent(missing_pct, accuracy = 0.01)),
    hjust = -0.1,
    size = 3
  ) +
  scale_x_continuous(
    labels = percent,
    expand = expansion(mult = c(0, 0.2))
  ) +
  labs(
    title = "Missing values are concentrated in selected vehicle-related fields",
    subtitle = "The affected fields are retained because they are still useful for segment-level analysis",
    x = "Percentage of missing records",
    y = NULL
  )
```

The missing value check shows that only four variables have a small portion of missing values: `fuel_type`, `vehicle_value`, `vehicle_age` and `age_driving_licence`.

For this exercise, missing values will not be removed because the main profitability and volatility measures, such as premium, claims, incurred amount and exposure, are still available. Removing rows with minor missing values may unnecessarily distort the portfolio-level analysis.

## Check duplicates
Duplicate records are checked to ensure that there are no repeated rows. If exact duplicates are found, they should be removed because they may represent repeated transaction records. However, repeated `insured_id` values alone should not be treated as duplicates, as the same customer may have multiple policy records across years or policy types.

```{r}
#| code-fold: true
duplicate_records <- sum(duplicated(motor_raw))
duplicate_pct <- duplicate_records / nrow(motor_raw)

duplicate_tbl <- data.frame(
  Check = "Exact duplicate rows across all variables",
  Duplicate_Records = comma(duplicate_records),
  Duplicate_Percentage = percent(duplicate_pct, accuracy = 0.01)
)

knitr::kable(
  duplicate_tbl,
  caption = "Duplicate records check"
)
```
:::

### 4.4 Data Preparation
#### 4.4.1 Deriving analytical variables
The raw dataset includes premium, incurred claims, number of claims and exposure. To investigate profitability and volatility, derived fields are created:

- `underwriting_result` = total premium minus total incurred claims,
- `loss_ratio` = total incurred claims divided by total premium,
- `claim_flag` = whether the record has at least one claim, and
- `loss_ratio_capped` = loss ratio capped at the 99th percentile for visualisation readability.

These derived fields are sufficient for the study objectives. `underwriting_result` and `loss_ratio` are used to assess profitability, while `claim_flag` and `loss_ratio_capped` help examine claim occurrence and volatility. Other variables are retained in their original form to keep the analysis simple and interpretable.

```{r}
motor_prep <- mutate(
  motor_raw,
  year = as.factor(year),
  policy_type = as.factor(policy_type),
  policy_status = as.factor(policy_status),
  business_type = as.factor(business_type),
  payment_frequency = as.factor(payment_frequency),
  underwriting_result = total_premium - total_incurred,
  loss_ratio = if_else(total_premium > 0, total_incurred / total_premium, NA_real_),
  claim_flag = total_claims > 0
)

motor_prep <- filter(
  motor_prep,
  total_premium > 0,
  total_exposure > 0
)

loss_ratio_cap_value <- quantile(motor_prep$loss_ratio, 0.99, na.rm = TRUE)

motor <- mutate(
  motor_prep,
  loss_ratio_capped = pmin(loss_ratio, loss_ratio_cap_value)
)
```

::: {.callout-note}
`loss_ratio_capped` is used only for visualisation and non-parametric tests. Portfolio profitability calculations still use the original premium and incurred claim values.
:::

#### 4.4.2 Portfolio-level summary
```{r}
#| code-fold: true
portfolio_summary <- summarise(
  motor,
  valid_records = n(),
  unique_insured = n_distinct(insured_id),
  total_premium = sum(total_premium, na.rm = TRUE),
  total_incurred = sum(total_incurred, na.rm = TRUE),
  underwriting_result = sum(underwriting_result, na.rm = TRUE),
  loss_ratio = sum(total_incurred, na.rm = TRUE) / sum(total_premium, na.rm = TRUE),
  total_claims = sum(total_claims, na.rm = TRUE),
  claim_record_rate = mean(claim_flag, na.rm = TRUE),
  p99_loss_ratio = quantile(loss_ratio, 0.99, na.rm = TRUE)
)

portfolio_display <- data.frame(
  Metric = c(
    "Valid records",
    "Unique insured customers",
    "Total premium",
    "Total incurred claims",
    "Underwriting result",
    "Portfolio loss ratio",
    "Total claims",
    "Claim record rate",
    "99th percentile loss ratio"
  ),
  Value = c(
    comma(portfolio_summary$valid_records),
    comma(portfolio_summary$unique_insured),
    dollar(portfolio_summary$total_premium, prefix = "$"),
    dollar(portfolio_summary$total_incurred, prefix = "$"),
    dollar(portfolio_summary$underwriting_result, prefix = "$"),
    percent(portfolio_summary$loss_ratio, accuracy = 0.1),
    comma(portfolio_summary$total_claims),
    percent(portfolio_summary$claim_record_rate, accuracy = 0.1),
    percent(portfolio_summary$p99_loss_ratio, accuracy = 0.1)
  )
)

knitr::kable(
  portfolio_display,
  caption = "Portfolio-level performance summary"
)
```

The portfolio contains **354,091** transaction records after filtering for positive premium and exposure. These records belong to **185,677** unique customers, showing that some customers appear more than once for transactions or renewals.

The portfolio looks **profitable**. It collected about *$105.1 million* in premiums and incurred about *$73.9 million* in claims, giving a positive underwriting result of about *$31.2 million*. The portfolio loss ratio is **70.3%**, which is below 100%, so claims did not exceed premiums. Only **14.3%** of records had claims, which means most records did not make a claim. However, there were still **77,098** claims in total. This suggests that the volatility comes from policies with more expensive claims.

Overall, the portfolio is profitable overall, but further analysis is needed to find which variables are less profitable or more volatile.

## 5. Objective 1: Univariate Visual Analytics
This section applies suitable univariate visual analytics techniques to explore the distributions and characteristics of numerical and categorical variables.

### 5.1 Distribution of Key Financial Variables
Histograms are suitable for identifying skewness, zero inflation and long-tail behaviour in numerical variables.

```{r}
#| code-fold: true
p1 <- ggplot(motor, aes(x = total_premium)) +
  geom_histogram(bins = 45, fill = "#e75480", colour = "white") +
  scale_x_continuous(labels = label_number()) +
  labs(
    title = "Total premium",
    x = "Total premium",
    y = "Number of records"
  )

p2 <- ggplot(motor, aes(x = total_incurred)) +
  geom_histogram(bins = 45, fill = "#e75480", colour = "white") +
  scale_x_continuous(labels = label_number()) +
  labs(
    title = "Total incurred claims",
    x = "Total incurred claims",
    y = "Number of records"
  )

p3 <- ggplot(motor, aes(x = loss_ratio_capped)) +
  geom_histogram(bins = 45, fill = "#e75480", colour = "white") +
  scale_x_continuous(labels = label_number()) +
  labs(
    title = "Loss ratio, capped at 99th percentile",
    x = "Loss ratio",
    y = "Number of records"
  )

p4 <- ggplot(motor, aes(x = underwriting_result)) +
  geom_histogram(bins = 45, fill = "#e75480", colour = "white") +
  scale_x_continuous(labels = label_number()) +
  labs(
    title = "Underwriting result",
    x = "Underwriting result",
    y = "Number of records"
  )

p1 + p2 + p3 + p4
```

The distributions show that the financial variables are highly skewed. Most records have low premiums and little or no incurred claims, so the bars are heavily concentrated near zero. This means the portfolio is generally stable for most policies. However, the underwriting result shows a few large negative cases, suggesting that profitability can still be affected by a small number of high-claim records.

### 5.2 Distribution of Customer and Vehicle Characteristics
```{r}
#| code-fold: true

p5 <- ggplot(motor, aes(x = driver_age)) +
  geom_histogram(bins = 40, fill = "#e75480", colour = "white") +
  labs(
    title = "Driver age",
    x = "Driver age",
    y = "Number of records"
  )

p6 <- ggplot(motor, aes(x = vehicle_value)) +
  geom_histogram(bins = 40, fill = "#e75480", colour = "white") +
  scale_x_continuous(labels = label_number()) +
  labs(
    title = "Vehicle value",
    x = "Vehicle value",
    y = "Number of records"
  )

p5 + p6 
```

From the plot, we can say that most drivers are *middle-aged*, while vehicle value is highly skewed, meaning most vehicles have *lower values*, with only a small number of very high-value vehicles.

### 5.3 Distribution of Claim Count
Because claim count is a discrete numerical variable, a bar chart is more suitable than a histogram.

```{r}
#| code-fold: true
claims_tbl <- mutate(
  motor,
  claims_group = if_else(total_claims >= 4, "4+", as.character(total_claims))
)

claims_tbl <- count(claims_tbl, claims_group)

claims_tbl <- mutate(
  claims_tbl,
  claims_group = factor(claims_group, levels = c("0", "1", "2", "3", "4+")),
  pct = n / sum(n)
)

ggplot(claims_tbl, aes(x = claims_group, y = pct)) +
  geom_col(fill = "#e75480") +
  geom_text(
    aes(label = percent(pct, accuracy = 0.1)),
    vjust = -0.4,
    size = 3.5
  ) +
  scale_y_continuous(
    labels = percent,
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title = "Percentage of Claims",
    x = "Total claims per record",
    y = "Share of records"
  )
```

Most records did not make any claims, showing that claims are uncommon in the portfolio. This also explains portfolio risk is driven by the smaller number of records with claims.

### 5.4 Categorical portfolio composition
Bar charts are suitable for categorical variables because they show how portfolio volume is distributed across policy, status, business and payment frequency categories.

::: {.panel-tabset}
## Policy Type
```{r}
#| code-fold: true
policy_tbl <- count(motor, policy_type)
policy_tbl <- mutate(policy_tbl, pct = n / sum(n))

p_policy <- ggplot(policy_tbl, aes(x = n, y = fct_reorder(policy_type, n))) +
  geom_col(fill = "#e75480") +
  geom_text(aes(label = percent(pct, accuracy = 0.1)), hjust = -0.1, size = 3) +
  scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Policy type",
    x = "Number of records",
    y = NULL
  )
p_policy
```
**CC (Third-party liability)** is the dominant category, making up **57.8%** of records, followed by **COMP_E (comprehensive with excess)** at **25.8%**. The portfolio is mainly concentrated in these two product types.

## Policy Status
```{r}
#| code-fold: true
status_tbl <- count(motor, policy_status)
status_tbl <- mutate(status_tbl, pct = n / sum(n))

p_status <- ggplot(status_tbl, aes(x = n, y = fct_reorder(policy_status, n))) +
  geom_col(fill = "#e75480") +
  geom_text(aes(label = percent(pct, accuracy = 0.1)), hjust = -0.1, size = 3) +
  scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Policy status",
    x = "Number of records",
    y = NULL
  )
p_status
```
Most records are active policies, accounting for **89.2%**. Although cancelled policies only accounts for **10.8%**, they should still be checked because it may affect profitability.

## Business Type
```{r}
#| code-fold: true
business_tbl <- count(motor, business_type)
business_tbl <- mutate(business_tbl, pct = n / sum(n))

p_business <- ggplot(business_tbl, aes(x = n, y = fct_reorder(business_type, n))) +
  geom_col(fill = "#e75480") +
  geom_text(aes(label = percent(pct, accuracy = 0.1)), hjust = -0.1, size = 3) +
  scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Business type",
    x = "Number of records",
    y = NULL
  )
p_business
```
New business records make up **69.3%**. This suggests that the portfolio is more heavily driven by new policy transactions.

## Payment Frequency
```{r}
#| code-fold: true
payment_tbl <- count(motor, payment_frequency)
payment_tbl <- mutate(payment_tbl, pct = n / sum(n))

p_payment <- ggplot(payment_tbl, aes(x = n, y = fct_reorder(payment_frequency, n))) +
  geom_col(fill = "#e75480") +
  geom_text(aes(label = percent(pct, accuracy = 0.1)), hjust = -0.1, size = 3) +
  scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Payment frequency",
    x = "Number of records",
    y = NULL
  )
p_payment
```
**Annual** payment is the most common at **81.3%**. **Semi-annual** and **quarterly** payments are much smaller groups, but they maybe still useful for analysis because payment behaviour can be related to customer profile and profitability.
:::

## 6. Objective 2: Bivariate Visual Analytics
This section applies suitable bivariate visual analytics techniques to investigate factors influencing portfolio profitability and volatility. Statistical validation is included to check whether the visual differences are supported by formal tests.

### 6.1 Profitability and Volatility over Time
```{r}
#| code-fold: true
year_perf <- group_by(motor, year)

year_perf <- summarise(
  year_perf,
  total_premium = sum(total_premium, na.rm = TRUE),
  total_incurred = sum(total_incurred, na.rm = TRUE),
  loss_ratio = total_incurred / total_premium
)

ggplot(year_perf, aes(x = year, y = loss_ratio)) +
  geom_col(fill = "#e75480") +
  geom_text(
    aes(label = percent(loss_ratio, accuracy = 0.1)),
    vjust = -0.4,
    size = 3.5
  ) +
  scale_y_continuous(labels = percent, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Portfolio loss ratio by year",
    x = "Year",
    y = "Loss ratio"
  )
```
The portfolio remained profitable across all three years because the loss ratio stayed below 100%. However, the loss ratio increased in 2024, suggesting that profitability weakened compared with 2022 and 2023.

::: {.callout-note}
A higher loss ratio means weaker profitability.
:::

### 6.2 Policy Type and Profitability 
```{r}
#| code-fold: true
policy_perf <- group_by(motor, policy_type)

policy_perf <- summarise(
  policy_perf,
  records = n(),
  premium = sum(total_premium, na.rm = TRUE),
  incurred = sum(total_incurred, na.rm = TRUE),
  loss_ratio = incurred / premium
)

ggplot(policy_perf, aes(x = loss_ratio, y = reorder(policy_type, loss_ratio))) +
  geom_col(fill = "#e75480") +
  geom_text(
    aes(label = percent(loss_ratio, accuracy = 0.1)),
    hjust = -0.1,
    size = 3
  ) +
  scale_x_continuous(
    labels = percent,
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title = "Loss ratio by policy type",
    x = "Loss ratio",
    y = "Policy type"
  )
```

The plot shows variation in profitability across policy types. `COMP_N` *(comprehensive without excess)* has the highest loss ratio at **93.9%**, indicating low profitability, while  `TP` *(basic coverage)* has the lowest loss ratio at **54.3%**, making it the most profitable segment. Although all the policy types remain below 100% and profitable before expenses, `COMP_N` should be monitored closely.

### 6.3 Claim Rate by Payment Frequency
```{r}
#| code-fold: true

payment_claim <- group_by(motor, payment_frequency)

payment_claim <- summarise(
  payment_claim,
  records = n(),
  claim_rate = mean(claim_flag, na.rm = TRUE)
)

ggplot(payment_claim, aes(x = reorder(payment_frequency, claim_rate), y = claim_rate)) +
  geom_col(fill = "#e75480") +
  geom_text(
    aes(label = percent(claim_rate, accuracy = 0.1)),
    vjust = -0.4,
    size = 3.5
  ) +
  scale_y_continuous(
    labels = percent,
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title = "Claim rate by payment frequency",
    x = "Payment frequency",
    y = "Share of records with claims"
  )
```

This plot compares how often claims occur across different payment frequencies. Payment groups with **higher claim rates** may indicate **greater claim volatility** and should be reviewed further, even if they represent a smaller share of the portfolio.

### 6.4 Policy Status and Business Type

::: {.panel-tabset}

## Policy Status
```{r}
#| code-fold: true

status_perf <- group_by(motor, policy_status)

status_perf <- summarise(
  status_perf,
  records = n(),
  loss_ratio = sum(total_incurred, na.rm = TRUE) / sum(total_premium, na.rm = TRUE)
)

ggplot(status_perf, aes(x = loss_ratio, y = reorder(policy_status, loss_ratio))) +
  geom_col(fill = "#e75480") +
  geom_text(aes(label = percent(loss_ratio, accuracy = 0.1)), hjust = -0.1, size = 3) +
  scale_x_continuous(labels = percent, expand = expansion(mult = c(0, 0.2))) +
  labs(
    title = "Loss ratio by policy status",
    x = "Loss ratio",
    y = "Policy status"
  )
```

## Business Type

```{r}
#| code-fold: true

business_perf <- group_by(motor, business_type)

business_perf <- summarise(
  business_perf,
  records = n(),
  loss_ratio = sum(total_incurred, na.rm = TRUE) / sum(total_premium, na.rm = TRUE)
)

ggplot(business_perf, aes(x = loss_ratio, y = reorder(business_type, loss_ratio))) +
  geom_col(fill = "#e75480") +
  geom_text(aes(label = percent(loss_ratio, accuracy = 0.1)), hjust = -0.1, size = 3) +
  scale_x_continuous(labels = percent, expand = expansion(mult = c(0, 0.2))) +
  labs(
    title = "Loss ratio by business type",
    x = "Loss ratio",
    y = "Business type"
  )
```
:::

Policy status and business type show differences in profitability. Groups with higher loss ratios are less profitable because a larger share of their premium is used to cover incurred claims.

### 6.5 Bonus Score and Claim Risk
Bonus score is analysed against claim rate to examine whether different customer risk groups show different levels of claim occurrence.
```{r}
#| code-fold: true
bonus_claim <- group_by(motor, bonus_score)

bonus_claim <- summarise(
  bonus_claim,
  records = n(),
  claim_rate = mean(claim_flag, na.rm = TRUE),
  loss_ratio = sum(total_incurred, na.rm = TRUE) / sum(total_premium, na.rm = TRUE)
)

ggplot(bonus_claim, aes(x = bonus_score, y = claim_rate)) +
  geom_col(fill = "#e75480") +
  geom_text(aes(label = percent(claim_rate, accuracy = 0.1)), vjust = -0.4) +
  scale_y_continuous(labels = percent, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Claim rate by bonus score",
    x = "Bonus score",
    y = "Share of records with claims"
  )
```
The plot shows a clear relationship between bonus score and claim risk. Records with a favourable claims history (`G`) have the lowest claim rate at **14.2%**, followed by neutral (`N`) at **17.6%**. while poor profiles (`B`) have the highest claim rate at **21.9%**. This indicates that bonus score is a meaningful indicator of volatility.

```{r, echo=FALSE}
bonus_table <- data.frame(
  `Bonus Score` = bonus_claim$bonus_score,
  Records = comma(bonus_claim$records),
  `Claim Rate` = percent(bonus_claim$claim_rate, accuracy = 0.1),
  `Loss Ratio` = percent(bonus_claim$loss_ratio, accuracy = 0.1)
)

knitr::kable(
  bonus_table,
  caption = "Claim risk and loss ratio by bonus score"
)
```

Higher claim rates show greater risk, while the table adds loss ratio to show how this translates into profitability.

## 7. Statistical Validation
The bivariate plots suggest that both profitability and claim risk differ across several portfolio segments. To support these observations, statistical tests are applied. Since loss ratio is highly skewed, **non-parametric tests** are used for profitability analysis. For claim occurrence, **Chi-square tests** are used as the outcome is categorical.

### 7.1 Validation of Profitability Differences
**Kruskal–Wallis tests** are used for variables with more than two categories (eg. year, policy type), while **Wilcoxon tests** are applied to variables with two groups (eg. policy status, business type).

```{r, echo=FALSE}
year_test <- kruskal.test(loss_ratio ~ year, data = motor)
policy_type_test <- kruskal.test(loss_ratio ~ policy_type, data = motor)
policy_status_test <- wilcox.test(loss_ratio ~ policy_status, data = motor)
business_type_test <- wilcox.test(loss_ratio ~ business_type, data = motor)

profitability_tests <- data.frame(
  Analysis = c(
    "Loss ratio by year",
    "Loss ratio by policy type",
    "Loss ratio by policy status",
    "Loss ratio by business type"
  ),
  Test = c(
    "Kruskal-Wallis test",
    "Kruskal-Wallis test",
    "Wilcoxon rank-sum test",
    "Wilcoxon rank-sum test"
  ),
  P_Value = c(
    year_test$p.value,
    policy_type_test$p.value,
    policy_status_test$p.value,
    business_type_test$p.value
  )
)

profitability_tests$P_Value <- format.pval(
  profitability_tests$P_Value,
  digits = 3,
  eps = 0.001
)

knitr::kable(
  profitability_tests,
  caption = "Statistical validation of profitability differences"
)
```

The profitability tests check whether the loss ratio differs across the selected groups. A small p-value shows that the differences observed are **statistically significant**, meaning that profitability is not evenly distributed across these segments.

### 7.2 Validation of Claim Risk Differences
**Chi-square tests** are used to validate whether claim occurrence is associated with payment frequency and bonus score.

```{r, echo=FALSE}
volatility_tests <- data.frame(
  Analysis = c(
    "Claim occurrence by payment frequency",
    "Claim occurrence by bonus score"
  ),
  Test = "Chi-square test",
  P_Value = format.pval(
    c(
      chisq.test(table(motor$payment_frequency, motor$claim_flag))$p.value,
      chisq.test(table(motor$bonus_score, motor$claim_flag))$p.value
    ),
    digits = 3,
    eps = 0.001
  )
)

knitr::kable(
  volatility_tests,
  caption = "Statistical validation of volatility differences"
)
```

The tests check whether differences in claim occurrence across these groups are **statistically significant**. A small p-value shows that claim risk is not randomly distributed, but associated with these factors.

### 7.3 Interpretation of Validation Results

Overall, the statistical tests support the bivariate visual analysis. Profitability differs across year, policy type, policy status and business type, while claim occurrence differs across payment frequency and bonus score. However, because the dataset is very large, statistical significance should be interpreted together with the charts and business context. The results show association, not causation.











