---
title: "Hands-on Exercise 3a: Programming Interactive Data Visualisation with R"
author: "Htet Kay Khaing"
date: "April 29, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 3.1 Learning Outcome

In this hands-on exercise, we will learn how to create interactive data visualisation by using functions provided by **ggiraph** and **plotlyr** packages.

## 3.2 Getting Started

First, write a code chunk to check, install and launch the following R packages:

-   [**ggiraph**](https://davidgohel.github.io/ggiraph/) for making ggplot graphics interactive.
-   [**plotly**](https://plotly.com/r/), R library for plotting interactive statistical graphs.
-   [**DT**](https://rstudio.github.io/DT/) provides an R interface to the JavaScript library [DataTables]() that create interactive table on html page.
-   [**tidyverse**](https://www.tidyverse.org/), a family of modern R packages specially designed to support data science, analysis and communication task including creating static statistical graphs.
-   [**patchwork**](https://patchwork.data-imaginist.com/) for combining multiple ggplot2 graphs into one figure.

The code chunk below will be used to accomplish the task.

```{r}
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse) 
```

## 3.3 Importing Data

In this section, *Exam_data.csv* provided will be used. Using [*read_csv()*](https://readr.tidyverse.org/reference/read_delim.html) of **readr** package, import *Exam_data.csv* into R.

The code chunk below `read_csv()` of **readr** package is used to import `Exam_data.csv` data file into R and save it as an tibble data frame called `exam_data`.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## 3.4 Interactive Data Visualisation - ggiraph methods

[**ggiraph**](https://davidgohel.github.io/ggiraph/index.html) is an htmlwidget and a ggplot2 extension. It allows ggplot graphics to be interactive.

Interactive is made with [**ggplot geometries**](https://davidgohel.github.io/ggiraph/reference/index.html#section-interactive-geometries) that can understand three arguments:

-   **Tooltip**: a column of data-sets that contain tooltips to be displayed when the mouse is over elements.
-   **Onclick**: a column of data-sets that contain a JavaScript function to be executed when elements are clicked.
-   **Data_id**: a column of data-sets that contain an id to be associated with elements.

If it used within a shiny application, elements associated with an id (data_id) can be selected and manipulated on client and server sides. Refer to this [article](https://www.ardata.fr/ggiraph-book/selections.html) for more detail explanation.

### Tooltip effect with *tooltip* aesthetic

Below shows a typical code chunk to plot an interactive statistical graph by using **ggiraph** package. Notice that the code chunk consists of two parts. First, an ggplot object will be created. Next, [`girafe()`](https://davidgohel.github.io/ggiraph/reference/girafe.html) of **ggiraph** will be used to create an interactive svg object.

```{r}
#| output: false
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE, 
    binwidth = 1, 
    method = "histodot") +
  scale_y_continuous(NULL, 
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```

Notice that two steps are involved. First, an interactive version of ggplot2 geom (i.e. [`geom_dotplot_interactive()`](https://davidgohel.github.io/ggiraph/reference/geom_dotplot_interactive.html)) will be used to create the basic graph. Then, [`girafe()`](https://davidgohel.github.io/ggiraph/reference/girafe.html) will be used to generate an svg object to be displayed on an html page.

## 3.5 Interactivity
By hovering the mouse pointer on an data point of interest, the student’s ID will be displayed.

```{r, echo=FALSE}
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```

### 3.5.1 Displaying multiple information on tooltip
The content of the tooltip can be customised by including a list object as shown in the code chunk below.

```{r}
#| output: false
exam_data$tooltip <- c(paste0(     
  "Name = ", exam_data$ID,         
  "\n Class = ", exam_data$CLASS)) 

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip), 
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot") +
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 8,
  height_svg = 8*0.618
)
```

The first three lines of codes in the code chunk create a new field called *tooltip*. At the same time, it populates text in ID and CLASS fields into the newly created field. Next, this newly created field is used as tooltip field as shown in the code of line 7.

## 3.6 Interactivity
By hovering the mouse pointer on an data point of interest, the student’s ID and Class will be displayed.

```{r, echo=FALSE}
girafe(
  ggobj = p,
  width_svg = 8,
  height_svg = 8*0.618
)
```

### 3.6.1 Customising Tooltip style
Code chunk below uses [opts_tooltip()](https://davidgohel.github.io/ggiraph/reference/opts_tooltip.html) of ggiraph to customize tooltip rendering by add css declarations.

```{r}
#| output: false
tooltip_css <- "background-color:white; #<<
font-style:bold; color:black;" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

Notice that the background colour of the tooltip is black and the font colour is white and bold.

```{r, echo=FALSE}
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)
```

- Refer to [Customizing girafe objects](https://www.ardata.fr/ggiraph-book/customize.html) to learn more about how to customise ggiraph objects.

### 3.6.2 Displaying statistics on tooltip
Code chunk below shows an advanced way to customise tooltip. In this example, a function is used to compute 90% confident interval of the mean. The derived statistics are then displayed in the tooltip.

```{r}
tooltip <- function(y, ymax, accuracy = .01) {
  mean <- scales::number(y, accuracy = accuracy)
  sem <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean maths scores:", mean, "+/-", sem)
}

gg_point <- ggplot(data=exam_data, 
                   aes(x = RACE),
) +
  stat_summary(aes(y = MATHS, 
                   tooltip = after_stat(  
                     tooltip(y, ymax))),  
    fun.data = "mean_se", 
    geom = GeomInteractiveCol,  
    fill = "light blue"
  ) +
  stat_summary(aes(y = MATHS),
    fun.data = mean_se,
    geom = "errorbar", width = 0.2, size = 0.2
  )

girafe(ggobj = gg_point,
       width_svg = 8,
       height_svg = 8*0.618)
```

### 3.6.3 Hover effect with *data_id* aesthetic
Code chunk below shows the second interactive feature of ggiraph, namely `data_id`.

```{r}
#| output: false
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(           
    aes(data_id = CLASS),             
    stackgroups = TRUE,               
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618                      
)                                        
```

Interactivity: Elements associated with a *data_id* (i.e CLASS) will be highlighted upon mouse over.

```{r, echo=FALSE}
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618                      
)
```

Note that the default value of the hover css is *hover_css = “fill:orange;”*.

### 3.6.4 Styling hover effect
In the code chunk below, css codes are used to change the highlighting effect.

```{r}
#| output: false
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```

Interactivity: Elements associated with a *data_id* (i.e CLASS) will be highlighted upon mouse over.

```{r, echo=FALSE}
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)
```

Note: Different from previous example, in this example the ccs customisation request are encoded directly.

### 3.6.5 Combining tooltip and hover effect
There are time that we want to combine tooltip and hover effect on the interactive statistical graph as shown in the code chunk below.

```{r}
#| output: false
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = CLASS, 
        data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```

Interactivity: Elements associated with a *data_id* (i.e CLASS) will be highlighted upon mouse over. At the same time, the tooltip will show the CLASS.

```{r, echo=FALSE}
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)
```

### 3.6.6 Click effect with onclick
`onclick` argument of ggiraph provides hotlink interactivity on the web.

The code chunk below shown an example of `onclick`.

```{r}
#| output: false
exam_data$onclick <- sprintf("window.open(\"%s%s\")",
"https://www.moe.gov.sg/schoolfinder?journey=Primary%20school",
as.character(exam_data$ID))

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(onclick = onclick),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618)                                        
```

Interactivity: Web document link with a data object will be displayed on the web browser upon mouse click.

```{r, echo=FALSE}
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618) 
```

::: {.callout-warning}
Note that click actions must be a string column in the dataset containing valid javascript instructions.
:::

### 3.6.7 Coordinated Multiple Views with ggiraph
Coordinated multiple views methods has been implemented in the data visualisation below.

```{r, echo=FALSE}
p1 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +  
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

p2 <- ggplot(data=exam_data, 
       aes(x = ENGLISH)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") + 
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

girafe(code = print(p1 + p2), 
       width_svg = 6,
       height_svg = 3,
       options = list(
         opts_hover(css = "fill: #202020;"),
         opts_hover_inv(css = "opacity:0.2;")
         )
       ) 
```

Notice that when a data point of one of the dotplot is selected, the corresponding data point ID on the second data visualisation will be highlighted too.

In order to build a coordinated multiple views as shown in the example above, the following programming strategy will be used:

1. Appropriate interactive functions of **ggiraph** will be used to create the multiple views.
2. *patchwork* function of [patchwork](https://patchwork.data-imaginist.com/) package will be used inside girafe function to create the interactive coordinated multiple views.

```{r}
#| output: false
p1 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +  
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

p2 <- ggplot(data=exam_data, 
       aes(x = ENGLISH)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") + 
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

girafe(code = print(p1 + p2), 
       width_svg = 6,
       height_svg = 3,
       options = list(
         opts_hover(css = "fill: #202020;"),
         opts_hover_inv(css = "opacity:0.2;")
         )
       ) 
```

The *data_id* aesthetic is critical to link observations between plots and the tooltip aesthetic is optional but nice to have when mouse over a point.









