---
title: "Hands-on Exercise 7b: Visualising and Analysing Text Data with R: tidytext methods"
author: "Htet Kay Khaing"
date: "May 31, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 7.1 Learning Outcome
In this chapter, we will visualise and analyse text data using R.

By the end of this chapter, we will be able to:

- understand tidytext framework for processing, analysing and visualising text data,
- write function for importing multiple files into R,
- combine multiple files into a single data frame,
- clean and wrangle text data by using tidyverse approach,
- visualise words with Word Cloud,
- compute term frequency–inverse document frequency (TF-IDF) using tidytext method, and
- visualising texts and terms relationship.

## 7.2 Getting Started
### 7.2.1 Installing and launching R packages
In this chapter, the following R packages for handling, processing, wrangling, analysing and visualising text data will be used:

- tidytext, tidyverse (mainly readr, purrr, stringr, ggplot2)
- widyr,
- wordcloud and ggwordcloud,
- textplot (required igraph, tidygraph and ggraph, )
- DT,
- lubridate and hms.

The code chunk:
```{r}
pacman::p_load(tidytext, widyr, wordcloud, DT, ggwordcloud, textplot, lubridate, hms,
tidyverse, tidygraph, ggraph, igraph)
```

## 7.3 Importing Multiple Text Files from Multiple Folders
### 7.3.1 Creating a folder list
```{r}
news20 <- "chap29/data/20news/"
```

### 7.3.2 Define a function to read all files from a folder into a data frame
```{r}
read_folder <- function(infolder) {
  tibble(file = dir(infolder, 
                    full.names = TRUE)) %>%
    mutate(text = map(file, 
                      read_lines)) %>%
    transmute(id = basename(file), 
              text) %>%
    unnest(text)
}
```

## 7.4 Importing Multiple Text Files from Multiple Folders
### 7.4.1 Reading in all the messages from the 20news folder
```{r}
raw_text <- tibble(folder = 
                     dir(news20, 
                         full.names = TRUE)) %>%
  mutate(folder_out = map(folder, 
                          read_folder)) %>%
  unnest(cols = c(folder_out)) %>%
  transmute(newsgroup = basename(folder), 
            id, text)
write_rds(raw_text, "data/rds/news20.rds")
```

::: {.callout-note title="Things to learn from the code chunk above"}

- [`read_lines()`](https://readr.tidyverse.org/reference/read_lines.html) of {**readr**](https://readr.tidyverse.org/) package is used to read up to n_max lines from a file.
- [`map()`](https://purrr.tidyverse.org/reference/map.html) of [**purrr**](https://purrr.tidyverse.org/) package is used to transform their input by applying a function to each element of a list and returning an object of the same length as the input.
- [`unnest()`](https://tidyr.tidyverse.org/reference/nest.html) of **dplyr** package is used to flatten a list-column of data frames back out into regular columns.
- [`mutate()`](https://dplyr.tidyverse.org/reference/mutate.html) of **dplyr** is used to add new variables and preserves existing ones;
- [`transmute()`](https://dplyr.tidyverse.org/reference/mutate.html) of **dplyr** is used to add new variables and drops existing ones.
- [`read_rds()`](https://readr.tidyverse.org/reference/read_rds.html) is used to save the extracted and combined data frame as rds file for future use.

:::

## 7.5 Initial EDA
Figure below shows the frequency of messages by newsgroup.



The code chunk:
```{r}
raw_text <- news20
raw_text %>%
  group_by(newsgroup) %>%
  summarize(messages = n_distinct(id)) %>%
  ggplot(aes(messages, newsgroup)) +
  geom_col(fill = "lightblue") +
  labs(y = NULL)
```
